---
title: "Link TDR data to environment"
author: "Benjamin Dupuis"
date: "`r Sys.Date()`"
format:
  html:
    self-contained: true
    code-fold: true
---

This script associates environmental and individual parameters.

```{r preambule, echo = F, output = F}
rm(list = ls())
gc()

library(tidyverse); library(ggpubr)
library(viridis)

theme_set(theme_bw())
```

First we load TDR and environmental data

```{r load tdr, echo = FALSE, message = FALSE}
ade_deploy <- read_csv("data/ade_deploy_1998-2024_final_wo_env.csv")
env_tibble <- read.csv("/media/bdupuis/1091_PhD_BD1/Data/Environmental data/env_data_copernicus.csv")

#adjust dates settings
env_tibble$date <- ymd(env_tibble$date)
env_tibble$season <- ifelse(month(env_tibble$date) <= 2,
                            year(env_tibble$date) - 1,
                            year(env_tibble$date))
env_tibble$year <- as.factor(year(env_tibble$date))
env_tibble$dummy_date <- env_tibble$date
year(env_tibble$dummy_date) <- 1904
year(env_tibble$dummy_date[which(month(env_tibble$dummy_date) <= 2)]) <- 1905

#remove feb 29th bug
env_tibble <- env_tibble %>% filter(!is.na(dummy_date))

env_tibble$days_since_start_season <- env_tibble$dummy_date - ymd("1904-11-01")

#rename some cols for clarity
env_tibble <- env_tibble %>%
  rename(mean_sic = sic)

glimpse(env_tibble)
```

Let's quickly look at the environmental variables.

First sea-ice variables

```{r plot env ice, warning = FALSE}
ggplot(data = env_tibble, aes(x = dummy_date, y = mean_sic, col = as.factor(season))) + 
  geom_line() +
  scale_color_viridis(discrete = T) + 
  labs(x = "Date", y = "SIC", col = "Breeding season") +  
  scale_x_date(date_labels = "%b", limits = c(ymd("1904-10-01", ymd("1905-02-28")))) +
  theme(legend.position = "right") +
  facet_wrap(~spatial_scale)

ggplot(data = env_tibble, aes(x = dummy_date, y = frac_ice, col = as.factor(season))) + 
  geom_line() +
  scale_color_viridis(discrete = T) + 
  labs(x = "Date", y = "Sea-ice fraction (>15%)", col = "Breeding season") +  
  scale_x_date(date_labels = "%b", limits = c(ymd("1904-10-01", ymd("1905-02-28")))) +
  theme(legend.position = "right") +
  facet_wrap(~spatial_scale)

ggplot(data = env_tibble, aes(x = dummy_date, y = frac_fast, col = as.factor(season))) + 
  geom_line() +
  scale_color_viridis(discrete = T) + 
  labs(x = "Date", y = "Fast ice fraction (>80%)", col = "Breeding season") +  
  scale_x_date(date_labels = "%b", limits = c(ymd("1904-10-01", ymd("1905-02-28")))) +
  theme(legend.position = "right") +
  facet_wrap(~spatial_scale)

ggplot(data = env_tibble, aes(x = dummy_date, y = frac_pack, col = as.factor(season))) + 
  geom_line() +
  scale_color_viridis(discrete = T) + 
  labs(x = "Date", y = "Pack ice fraction (between 15 and 80 %)", col = "Breeding season") +  
  scale_x_date(date_labels = "%b", limits = c(ymd("1904-10-01", ymd("1905-02-28")))) +
  theme(legend.position = "right") +
  facet_wrap(~spatial_scale)

retreat_tibble <- env_tibble %>% 
  filter(spatial_scale == "S1", mean_sic < 0.20, dummy_date > ymd("1904-10-01")) %>% 
  group_by(season) %>% 
  summarise(ice_retreat_date = min(dummy_date),
            ice_retreat_day = min(days_since_start_season))

retreat_tibble %>% 
  ggplot(aes(x = season, y = ice_retreat_date)) +
  geom_point(size = 2) +
  geom_line(linewidth = 1) +
  labs(x = "Season", y = "Sea-ice retreat date", title = "Earliest date when sea-ice concentration falls below 20%", subtitle = "Measured over Adélie penguins wintering area") +
  labs_pubr() +
  theme_pubr()
ggsave(filename = "figures/env_data/sic_retreat_plot.png", width = 9, height = 6, dpi = 500)
#smoothed line 

library(zoo)
env_tibble_smooth <- env_tibble %>% filter(spatial_scale == "S1")
env_tibble_smooth$sic_smooth <- rollmean(env_tibble_smooth$mean_sic, k = 7, align = "left", fill = c(NA, NA, NA))
sic_plot <- ggplot(data = env_tibble_smooth, aes(x = dummy_date, y = sic_smooth, col = as.factor(season))) + 
  geom_line(linewidth = 1) +
  scale_color_viridis(discrete = T, option = "D") + 
  labs_pubr() +
  theme_pubr() +
  labs(x = "Date", y = "SIC", col = "Breeding season") +  
  scale_x_date(date_labels = "%b", limits = c(ymd("1904-10-01", ymd("1905-02-28")))) +
  theme(legend.position = "top") +
  guides(color=guide_legend(nrow=3, byrow=TRUE))

ggsave(plot = sic_plot, filename = "figures/env_data/sic_smooth_plot.png", width = 9, height = 6, dpi = 500)
```

Now we add sea-ice retreat dates (cf. youngflesh et al. 2017 in Ecology) and phytoplantkon bloom intensity and temporality

```{r}
sic_cr <- env_tibble %>% filter(spatial_scale == "S3" & dummy_date >= dmy("20-12-1904") & dummy_date <= dmy("20-01-1905")) %>% 
  group_by(season) %>% 
  summarise(cr_sic_S3 = mean(mean_sic),
            cr_pack_S3 = mean(frac_pack),
            cr_fast_S3 = mean(frac_fast))

#winter here is july + august
sic_winter <- env_tibble %>% filter(spatial_scale == "S1" & dummy_date >= dmy("01-07-1904") & dummy_date <= dmy("31-08-1904")) %>% 
  group_by(season) %>% 
  summarise(winter_sic_S1 = mean(mean_sic),
            winter_pack_S1 = mean(frac_pack),
            winter_fast_S1 = mean(frac_fast))

retreat_tibble <- left_join(left_join(retreat_tibble, sic_winter), sic_cr)
```

From that we can extract several variables that we will use in future models (average, sd, min, max etc.)

```{r}
#trip level
ade_deploy$sic_trip <- NA
ade_deploy$pack_trip <- NA
ade_deploy$fast_trip <- NA

for(i in 1:nrow(ade_deploy)){
  #trip lvl
  start_trip <- date(ade_deploy$start_trip[i])
  end_trip <- date(ade_deploy$end_trip[i])
  trip_dates <- seq.Date(from = start_trip, to = end_trip, by = 1)
  sub_env_data <- env_tibble[which(env_tibble$date %in% trip_dates),] %>% filter(spatial_scale == "S3")
  
  ade_deploy$sic_trip[i] <- mean(sub_env_data$mean_sic, na.rm = T)
  ade_deploy$pack_trip[i] <- mean(sub_env_data$frac_pack, na.rm = T)
  ade_deploy$fast_trip[i] <- mean(sub_env_data$frac_fast, na.rm = T)
}

#Add season env
ade_deploy <- left_join(ade_deploy, retreat_tibble, by = join_by(year == season))

ade_deploy <- ade_deploy %>% dplyr::select(season, year, tag_id,
                             bm_init, flipper_length_mm,
                             logger_type, mass_end_kg , logger_id, colony, n_chick_equip,
                             start_trip, end_trip, trip_duration_days,
                             mean_depth_foraging, mean_dive_duration,
                             daily_dive_duration_hrs,
                             dee, daily_time_foraging_sec,
                             sic_trip, pack_trip, fast_trip,
                             ice_retreat_day,
                             winter_sic_S1, winter_pack_S1, winter_fast_S1,
                             cr_sic_S3, cr_pack_S3, cr_fast_S3)

ade_deploy$relative_departure_day <- date(ade_deploy$start_trip) - dmy(paste0("01-11-",ade_deploy$year))
ade_deploy$n_day_since_ice_retreat <- ade_deploy$relative_departure_day - ade_deploy$ice_retreat_day

#Add date relative to hatching peak + BS
bs_tibble <- read.csv("data/breeding_success_ade.csv") %>% 
  dplyr::select(year, FS_janv, hatching_peak, prop_succeed_incub)

ade_deploy <- left_join(ade_deploy, bs_tibble)
ade_deploy$departure_rel_hatch_peak <- date(ade_deploy$start_trip) - dmy(ade_deploy$hatching_peak)

#Add body condition
lm_body_condition <- lm(bm_init ~ flipper_length_mm,
   data = ade_deploy)

ade_deploy <- ade_deploy %>%
  mutate(body_condition = resid(lm_body_condition))

mf_tibble <- read.csv("/media/bdupuis/1091_PhD_BD1/Data/Environmental data/mf_yearly_data.csv")

ade_deploy <- left_join(ade_deploy, mf_tibble)
write_csv(ade_deploy, "data/ade_deploy_1998-2024_w_env_bs.csv")
```

Prepare the csv's for NetLogo

```{r}
daily_ibm <- env_tibble %>%
  filter(spatial_scale == "S3") %>%
  select(date, season, mean_sic) %>% 
  rename(year = season)

#add hatching peak date
peak_tibble <- read.csv("data/hatching_peak_ade.csv")
daily_ibm <- left_join(daily_ibm, peak_tibble) %>% 
  mutate(hatching_peak = dmy(hatching_peak),
         days_since_hatching_peak = date - hatching_peak) %>% 
  filter(!is.na(hatching_peak), month(date) %in% c(12, 1, 2)) %>% 
  arrange(date) %>% 
  select(days_since_hatching_peak, date, year, mean_sic)

#yearly data
yearly_ibm <- retreat_tibble %>%
  rename(year = season) %>% 
  select(year, ice_retreat_day, winter_sic_S1) %>% 
  filter(!is.na(winter_sic_S1))

yearly_ibm <- left_join(yearly_ibm, mf_tibble %>% select(year, n_day_snow))

glimpse(daily_ibm)
glimpse(yearly_ibm)

write_csv(daily_ibm, "/media/bdupuis/1091_PhD_BD1/Research projects/AdeBM/data/daily_ibm_env_data.csv")
write_csv(yearly_ibm, "/media/bdupuis/1091_PhD_BD1/Research projects/AdeBM/data/yearly_ibm_env_data.csv")
```
